---
title: "DS_Assignment_06"
author: "Ankit"
date: "21/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) #loading,library
```

## 1. 

### 1.1 

```{r}
boston <- read.delim('boston.csv.bz2')  # loading dataset
head(boston)
```

```{r}
summary(boston) #summary
dim(boston)  #dimensions
sum(is.na(boston))  #missing
```

### 1.2

#### 1.2(a)

```{r}
ggplot(boston, aes(rm,medv)) +
  geom_point() + 
  geom_smooth(method="lm")  # scatter plot of rm vs medv
```

```{r}
ggplot(boston, aes(lstat, medv)) +
  geom_point() + 
  geom_smooth(method="lm")  # scatter plot of lstat vs medv
```

```{r}
ggplot(boston, aes(indus, medv)) +
  geom_point() + 
  geom_smooth(method="lm")  # scatter plot of indus vs medv
```

Comment the result: 

We can see from the the graph of rm, it is showing positive association with the medv whereas lstat and indus is showing  negatively association with the medv.  

Furthermoe, graph for indus is also showing a negative relationship and so is the model where indus is statistically sigificantly and positive relationship with medv.  

For lstat and rm, all the data points are very close to the regression line but for indus, the data points are scattered. 


### 1.2(b)

```{r}
summary(m <- lm(medv ~ rm, data=boston)) # model for rm variable
```

```{r}
summary(m <- lm(medv ~ lstat, data=boston))  # model for lstat variable 
```

```{r}
summary(m <- lm(medv ~ indus, data=boston))  #model for indus variable 
```

### 1.2(c)

For rm, it is statistically significant (p value is less than 0.05 which lies within 5% CI). Also, as we increase 1 unit of rm it will increase 9.102 units of medv. 

For lstat, it is statistically significant (p value is less than 0.05 which lies within 5% CI). As we increase 1 unit of lstat it will decrease  -0.95 units of medv.

For indus, it is statistically significant (p value is less than 0.05 which lies within 5% CI). As we increase 1 unit of indus it will decrease -0.64 units of medv.

### 1.2(d)

For rm, as the average number of rooms per dwelling increases, the cost of the apartment is also supposed to increase which suggests that people in the neighbourhood are wealthy as they are able to afford expensive apartemtns which suggests that the whole neighbourhood expensive (suggested by the graph also, as the number of room is inreasing the medv is also increasing). 

For lstat, the prices of the the houses would be low where the lower status poppulation is living which suggests that the neighbourhood is not wealthy moreover to a lower background. Another point is as they cannot afford expensive houses. This can be seen from the graph that as the lstat is increase medv is decreasing. 

For indus which is neighbourhood with mostly non retail business suggests us two things:

1. The neighbourhood has a lot of manufacturing factories, which suggests that there are no/less houses and as there are less houses the value of the houses are also very less(cheap) which is exactly what the graph suggests us as the non retail business increases the medv would decrease. 

2. If the neighbourhood doesn't have retail businesses, people would not want to live there as daily use items won't be available easaily which means that it will be difficult for the people in neighborhood to buy groceries and other stuff. So, why would anyone wants to live in such a place unless the person is not rich. Hence this suggests that the neighbourhood is with people who are not earning much so is the plot. 

### 1.3

Yes, the relationship suggested by the plot is related to the model which is statistically significant. As we can see from the the graph of rm, it is showing positive association with the medv and from the model also we can see that rm is statistically significant and in positive relationship with medv. 

Similarly for lstat, we can see from the graph that lstat is negatively associated with the medv and same is the case with the model. Model is statistically significant and negative relationship with medv. 

Furthermoe, graph for indus is also showing a negative relationship and so is the model where indus is statistically sigificantly and positive relationship with medv.  

### 1.4

T value computed by the formula -> estimate/standard error. 
The null hypotheses(H0) here is that mdev is not assocaited with rm. 

To calculate t critical value, we have to use t distribution table at 95% confidence interval and two tail test.
Our degree of freedom isi 504 and two tail value in 0.05, so out t critical is 1.962. 

### 1.5
```{r}
summary(m <- lm(medv ~ ., data=boston))
```

As we can see from the above summary, except two variables, all the other variables are statistically significant. Two variables are indus and age. 

Except indus and age, we will reject all the other variables. 

### 1.6

For rm, 
rm is statistically significant as we can see that its p value is less than 0.05 which lies in 5% confidence interval. 

For lstat, 
similarly for lstat, it is statistically significant as we can see that its p value is less than 0.05 which lies in 5% confidence interval. 

For indus, 
it is not statistically significant as we can see that its p value is more than 0.05 which doesn't lie in 5% confidence interval. 

### 1.7
If we compare the results from 2 and 5: 

For rm, t value in question 2 is 21.72 and in question 5 it is 9.116. They are different because of direct and indirect relationships between rm and medv. 
In simple regression, rm and medv are in direct relationship but there are other indirect relationships between other variables and medv and rm. Whereas in the case of question 5, becaue multiple variables are exerting direct relationships with the medv which is neglecting indirect relationships between multiple variables and rm. 

Similarly for lstat, in question 2 the t value is -24.53 and in question 5 - 10.347. The difference in the number is also because of the effect of direct and indirect relationship. 

Also, for indus because of the effect of indirect and direct realtion there is difference between question 2 (-12.41) and question 5 (0.334) and if we notice, indus is no more statistically significant in question 5. 


## 2. 

### 2.1
Number of evictions is positively associated with the 311 calls. More evictions see more 311 calls. One unit increase no. of evictions is 0.04 units of increase in calls.  

### 2.2
Yes it is. As we can see, no. of evictions are positively associated with the 311 calls. 

### 2.3
Poverty rate is negatively associated with the 311 calls. 1 unit increase in poverty rate leads to -0.14 unit decrease in 311 calls. 

### 2.4
As we can see from the table, race (%white) is negatively associated with the calls(logged) which means that one unit of race(%white) is -0.03 units decrease in calls (because it is negative it will eventually be decreasing).

### 2.5
Median age is positively associated with the 311 call which means that one unit increase in median age is 0.0067 unit increase in calls. As it was given in the table ***p<0.01, which means that the this is statistically significant at 1% level.

### 2.6
The value of housing density is -0.13. It means that one unit increase of housing density will leads to 0.13 units decrease in the 311 calls (because it is negative it will eventually be decreasing).

### 2.7
The omitted category for city is Austin, TX is an reference category. To compare it with the neighbourhood of Philadelphia, it has less calls by 0.56 unit. 


## 3. 

### 3.1
Number of evictions is positively associated with the 311 calls. More evictions see more 311 calls. One percent increase no. of evictions is 0.4 percent of increase in calls.  

### 3.2
Yes it is. As we can see, no. of evictions are positively associated with the 311 calls. 

### 3.3
Poverty rate is negatively associated with the 311 calls. 1 unit increase in poverty rate leads to -1.4% decrease in 311 calls. 

### 3.4
As we can see from the table, race (%white) is negatively associated with the calls(logged) which means that one unit larger race(%white) is -0.3% decrease in calls (because it is negative it will eventually be decreasing).

### 3.5
Meadian age is positively associated with the 311 call which means that one unit of median age is 0.067% increase in calls. As it was given in the table ***p<0.01, which means that the this is statistically significant at 1% level. 

### 3.6
The value of housing density is -0.13. It means that one percent increase in housing density will leads to 1.3% decreasae in the 311 calls (because it is negative it will eventually be decreasing). 

### 3.7
The omitted category for city is Austin, TX is an reference category. To compare it with the neighbourhood of Philadelphia, it has less calls by -0.56 units. 







































