---
title: "DS Assignment 01"
author: "Ankit"
date: "12/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic R Programming 

## 1. 

* Created a function **time_format** which takes an **argument time in HHMM format**. 
* It *extracts hour and min separately* by using **remainder and integer division** arithmetic operators. 
* Calculated *fraction of hour* by **dividing the minutes by 60 and multiplying by 100**.  
* Prints the *result* in **HH.fraction of hours** format. 
* Then takes time as an **input from user in HHMM format**. Converts it into *integer* and passes it as **an argument in function call**. 

```{r}
time_format <- function(time){   #function definition
  min <- time%%100 #Here using integer division I am getting quotient i.e. the hour part of the time
  hour <- time%/%100 #Here using remainder,  I am getting remainder i.e. the minute part of the time
  frac <- round(min/60*100,digit=0) 

  paste(hour,frac,sep = ".")
}

time_<- strtoi(readline(prompt="Enter time(in HHMM): ")) #taking input from he user and converting it into int(using stroi) to use it in the above defined function. 
time_format(time_)  #function calling
```

## 2.

* First, created an **empty vector** 
* Then, created an **random vector(v)** with positive and negative numbers
* Used **for loop** to parse through the random vector(v)
* Within for loop, used if condition to **check whether the number is positive or not**. 
* **Append the positive numbers** in the above created empty vector and **print** it outside the for loop.

```{r}
empty_v <- NULL  #defining an empty vector 

set.seed(1)
v <- sample(10, 20, replace=TRUE) - 5 #created a random vector with both positive and negative numbers. 

for(i in v){   #for loop starts
  if(i>0){     #if loop which checks positive values in vector v
    empty_v <- c(empty_v,i)  #appending positive values in the above created empty vector empty_v.
  }
}
print(empty_v)   #print empty_v with all the positive values of vector V. 
```

## 3. 

* First, created an **empty vector** 
* Then, created an **random vector(v1)** with positive and negative numbers
* Use **logical indexing to get only positive numbers** in the random vector(v1) and **assign them to the empty vector**.

```{r}
empty_v1 <- NULL  #defining an empty vector 

set.seed(1)
v1 <- sample(10, 20, replace=TRUE) - 5 #created a random vector with both positive and negative

empty_v1 <- v1[v1>0]
print(empty_v1)
```

## 4. 

* Assigned 3 vectors as it was given in the question. 
* Created test function to check whether any value in the vector is positive or negative.

```{r}
v1 <- 9
v2 <- c(1,2)
v3 <- c(2,3,-4)

test <- function(vector_){
  
  if(any(vector_<0)){
    print('Vector has some negative values!')
  } else{
    print("Vector doesn't have any negative values")
  }
} 

test(v1)
test(v2)
test(v3)
```


# 2.1 Data Exploration 

## 2.1.1

I think if I have a suitable dataset and an analysis tool I can answer the question. The data with suitable variables like type of attack, how deadly the attack is, date,country and total number of attacks would the most important variables. 

To get which country is dangerous, I think we can use total number of shark cases as compared to the population of the country. 
And as Australia has more population, cases would be higher there. 


## 2.2.2
```{r}

#import libraries 

library(readxl)  
library(tidyverse)
```
```{r, cache = TRUE}
shark_attack <- read_excel('GSAF5.xls')  #Read the excel file
```


```{r}
dim(shark_attack)  #dimensions of the data set, 25827 rows and 24 columns
```

## 2.1.3
```{r}
colnames(shark_attack)  #get the names of the columns.  
```
Columns which I feel are important to answer the question are: 
* Country(To compare Australia and South Africa we need to group by country)
* Fatal(To check how many fatal attacks has happened in our target countries -> will help understand how dangerous the country is)
* Type(To find the relationship between fatality and the type of attacks -> What kind of attacks are fatal)

# 2.2 Explore Data

## 2.2.1
```{r}
unique(shark_attack['Country']) %>% arrange(Country)  #Get all the countries 
```

## 2.2.2

After browsing the countries, some problems I can see are: 

Some countries have "?" in their name which is a tyoe if inconsistency
Some Country names are also the state names like Andaman and Nicobar
Old country names are there for Sri lanka some ceylon is there. 

## 2.2.3
```{r}
sapply(shark_attack['Year'],class)  # to get the data type of the variable Year
```
Year has character data type, it should be int/datetime


## 2.2.4
```{r}
shark_attack %>% filter(is.na(Year)) %>% nrow() #Get missing values of the variable Year
```
More than 70% of the observations are Null. So, we dont know the year of the attacks for more than 70% of the records we have. 
Another thing could be that system was flawed in old times that they couldn't get the date of the attack.


## 2.2.5

```{r}
shark_attack$Year <- as.numeric(shark_attack$Year) #Converting the variable Year to number to calculate median
```

```{r}
min(shark_attack$Year,na.rm=TRUE)
max(shark_attack$Year,na.rm=TRUE)
median(shark_attack$Year,na.rm=TRUE)
```


## 2.2.6
```{r}
shark_attack %>% filter(Year==0) %>% head()
shark_attack %>% filter(Year==0) %>% tail()
```

I have found 2 cases:

* The date is in BC where there is no year
* Another case is when we dont know the year but we know the approx date like "Before 1908" or "Between 1992-1998" 
* The above two observations bring me to my next point that year must be extracted from the date by apply condition "last 4 digits of the date" and in the cases as mentioned above we dont have a proper date format, so in these cases, they might have replaced these year with 0


## 2.2.7

Krater was found during excavations at Lacco Ameno and performed research in 1958 and published a paper for the same. 
```{r}
shark_attack %>% filter(Date == "Ca. 725 B.C.")
```


# 2.3 Clean Data


## 2.3.1

Variables I am choosing: Date, Year, Type, Country, Area, Activity, injury, Fatal (Y/N)

```{r}
colnames(shark_attack)
```

```{r}
shark_attack <- shark_attack %>% rename(fatal = 'Fatal (Y/N)')
```


```{r}
shark_attack_sub <- shark_attack %>% select(Year:Area,Activity, Injury, fatal)
```


## 2.3.2
40% of the data if we remove NAs are between these years therefore I have taken 20 years.
We are left with 2200 rows of data. 
```{r}
shark_attack_sub <- shark_attack_sub %>% filter(Year<=2019 & Year >=2000)
```

## 2.3.3
```{r}
unique(shark_attack_sub$fatal)
```
The ones I can figure out are: 

* N: Not Fatal
* Y/y: Yes, its Fatal
* Unknown: Don't know how deadly the attack was. 
* Y x 2: Looks like a data entry error of 'Y' but its an fatal attack
* Rest all of them looks like an data entry error

## 2.3.4
Converting Unknown, Nq, 2017, F to Null 
```{r}
shark_attack_sub <- shark_attack_sub %>%
  mutate(fatal = na_if(fatal, "UNKNOWN")) %>% mutate(fatal = na_if(fatal, "Nq")) %>% mutate(fatal = na_if(fatal, "F")) %>% mutate(fatal = na_if(fatal, 2017)) # using na_if to replace useless values to NA
```

Convert the rest into boolean
```{r}
shark_attack_sub <- shark_attack_sub %>% mutate(fatal=ifelse(fatal=="Y"|fatal=="Y x 2"|fatal=="y",TRUE,FALSE))  #using vectored ifelse to convert catogeries to Boolean. 
```


```{r}
unique(shark_attack_sub$fatal)
```
# 2.4 Australia or South Africa

## 2.4.1
```{r}
shark_attack_sub %>% filter(Country == "AUSTRALIA" | Country == "SOUTH AFRICA") %>% group_by(Country) %>% summarize(n=n(),fatalp=round(sum(fatal,na.rm = TRUE)*100/n(),2))
```

## 2.4.2
In my opinion, we can answer this question in two ways: 

* Let's say if we consider fatality% is our key variable to access danger then South Africa is dangerous because although the number of cases are low but the fatality% is almost double compared to Australia. So, it tells us that more dangerous shark attacks are happening in South Africa. 

* Other case is when we consider total cases as our key variable to access danger then Australia is dangerous because although the fatality are low, cases are 3 time compared to South Africa So, it tells us that although the cases are  dangerous shark attacks are happening in South Africa. 

The analysis and answer can be used for tourism industry. As these countries are dangerous because of shark attacks, it can help the people take their tourism choices. 

## 2.4.3

Ethical issues are: 

* Name and personal information of people should be anonymous.
* We should try to remain neutral and unbiased. Don’t let our personal preconceptions or opinions interfere with the data collection process.
* We should take people's consent before using theie informations.

```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```






























































































































































































































